---
title: "Cnat2-18_field_trials"
author: "J. Meyer"
date: "2026-07-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries

```{r}
library(dada2)
library(CoDaSeq)
library(ggplot2)
library(phyloseq)
library(vegan)
library(corncob)
library(knitr)
library(dplyr)
library(tibble)
library(reshape2)
library(cowplot)
writeLines(capture.output(sessionInfo()), "sessionInfo.txt")
```


## Quality-filter the sequencing reads and create Amplicon Sequence Variant (ASV) tables with DADA2
Put unjoined R1 and R2 fastq files, with adaptors and primers previously removed with cutadapt into a directory for DADA2. Here, our forward and reverse fastq filenames have format: SAMPLENAME_R1_cut.fastq.gz and SAMPLENAME_R2_cut.fastq.gz


```{r}
path <- "~/Documents/CNAT2-18_fieldtrials/cutadapt"
list.files(path)
fnFs <- sort(list.files(path, pattern="_R1_cut.fastq.gz", full.names = TRUE))
fnRs <- sort(list.files(path, pattern="_R2_cut.fastq.gz", full.names = TRUE))
sample.names <- sapply(strsplit(basename(fnFs), "_"), `[`, 1)
# Perform filtering and trimming
filt_path <- file.path(path, "filtered") 
filtFs <- file.path(filt_path, paste0(sample.names, "_F_filt.fastq.gz"))
filtRs <- file.path(filt_path, paste0(sample.names, "_R_filt.fastq.gz"))
out <- filterAndTrim(fnFs, filtFs, fnRs, filtRs, truncLen=c(150,150),
                     maxN=0, maxEE=c(2,2), truncQ=2, rm.phix=TRUE,
                     compress=TRUE, multithread=TRUE)
head(out)
# Learn the Error Rates, it TAKES TIME!
errF <- learnErrors(filtFs, multithread=TRUE)
errR <- learnErrors(filtRs, multithread=TRUE)
plotErrors(errF, nominalQ=TRUE)
# Dereplicate the filtered fastq files
derepFs <- derepFastq(filtFs, verbose=TRUE)
derepRs <- derepFastq(filtRs, verbose=TRUE)
names(derepFs) <- sample.names
names(derepRs) <- sample.names
# Infer the sequence variants in each sample
dadaFs <- dada(derepFs, err=errF, multithread=TRUE)
dadaRs <- dada(derepRs, err=errR, multithread=TRUE)
# Inspecting the dada-class object returned by dada:
dadaFs[[1]]
# Merge the denoised forward and reverse reads:
mergers <- mergePairs(dadaFs, derepFs, dadaRs, derepRs, verbose=TRUE)
# Inspect the merger data.frame from the first sample
head(mergers[[1]])
# Construct sequence table
seqtab <- makeSequenceTable(mergers)
dim(seqtab)
#Remove chimeric sequences:
seqtab.nochim <- removeBimeraDenovo(seqtab, method="consensus", multithread=TRUE, verbose=TRUE)
dim(seqtab.nochim)
sum(seqtab.nochim)/sum(seqtab)
# Inspect distribution of sequence lengths
table(nchar(getSequences(seqtab)))
# Track reads through the pipeline
getN <- function(x) sum(getUniques(x))
track <- cbind(out, sapply(dadaFs, getN), sapply(mergers, getN), rowSums(seqtab))
colnames(track) <- c("input", "filtered", "denoised", "merged", "tabled")
rownames(track) <- sample.names
head(track)
write.table(track, "dada_read_stats.txt",sep="\t",col.names=NA)
saveRDS(seqtab.nochim, "~/Documents/CNAT2-18_fieldtrials/seqtab.rds") 
```


## Assign taxonomy in DADA2

Make sure the taxonomy reference database is in your working directory. Keep the database file gzipped. Adjust path name below. This step is very time consuming.

When taxonomy assignment is complete, we will use base R and phyloseq to clean up the taxonomy table. First, we will replace NAs and empty cells with the lowest taxonomy classification available. Second, we will use phyloseq to remove reads that are classified as Eukaryotes or unclassified at the domain level (ie, we are keeping only Bacteria and Archaea because that is what our primers target).

```{r}
taxa <- assignTaxonomy(seqtab.nochim, "~/Documents/silva_nr99_v138.2_toGenus_trainset.fa.gz", multithread=TRUE)
# FIX the NAs in the taxa table
taxon <- as.data.frame(taxa,stringsAsFactors=FALSE)
taxon$Phylum[is.na(taxon$Phylum)] <- taxon$Kingdom[is.na(taxon$Phylum)]
taxon$Class[is.na(taxon$Class)] <- taxon$Phylum[is.na(taxon$Class)]
taxon$Order[is.na(taxon$Order)] <- taxon$Class[is.na(taxon$Order)]
taxon$Family[is.na(taxon$Family)] <- taxon$Order[is.na(taxon$Family)]
taxon$Genus[is.na(taxon$Genus)] <- taxon$Family[is.na(taxon$Genus)]
write.table(taxon,"silva_taxa_table.txt",sep="\t",col.names=NA)
write.table(seqtab.nochim, "silva_otu_table.txt",sep="\t",col.names=NA)

# Create phyloseq object from otu and taxonomy tables from dada2, along with the sample metadata.
otu <- read.table("silva_otu_table.txt",sep="\t",header=TRUE, row.names=1)
taxon <- read.table("silva_taxa_table.txt",sep="\t",header=TRUE,row.names=1)
samples<-read.table("metadata.txt",sep="\t",header=T,row.names=1)
OTU = otu_table(otu, taxa_are_rows=FALSE)
taxon<-as.matrix(taxon)
TAX = tax_table(taxon)
sampledata = sample_data(samples)
ps <- phyloseq(otu_table(otu, taxa_are_rows=FALSE), 
               sample_data(samples), 
               tax_table(taxon))
ps # 1351 taxa and 34 samples

# remove chloroplasts and mitochondria and Eukaryota
get_taxa_unique(ps, "Family") # 232
get_taxa_unique(ps, "Order") # 140
get_taxa_unique(ps, "Kingdom") # 2
ps <- subset_taxa(ps, Family !="Mitochondria")
ps <- subset_taxa(ps, Order !="Chloroplast")
ps <- subset_taxa(ps, Kingdom !="Eukaryota")
ps <- subset_taxa(ps, Kingdom !="NA")
get_taxa_unique(ps, "Family") # 230
get_taxa_unique(ps, "Order") # 139
get_taxa_unique(ps, "Kingdom") # 2
ps # 1321 taxa and 34 samples

# Now export cleaned otu and taxa tables from phyloseq for future reference
otu = as(otu_table(ps), "matrix")
taxon = as(tax_table(ps), "matrix")
metadata = as(sample_data(ps), "matrix")
write.table(otu,"silva_nochloronomito_otu_table.txt",sep="\t",col.names=NA)
write.table(taxon,"silva_nochloronomito_taxa_table.txt",sep="\t",col.names=NA)

# explore data
length(get_taxa_unique(ps, "Genus")) # 405 genera
length(get_taxa_unique(ps, "Family")) # 230 families
length(get_taxa_unique(ps, "Order")) # 139 orders
length(get_taxa_unique(ps, "Class")) # 61 classes
length(get_taxa_unique(ps, "Phylum")) # 31 phyla

ps5<-filter_taxa(ps, function(x) mean(x) >5, TRUE)
ntaxa(ps5) # 144 - this retains 11% of original ASVs
get_taxa_unique(ps5, "Genus") # 62 genera

ps2<-filter_taxa(ps, function(x) mean(x) >2, TRUE)
ntaxa(ps2) # 232, retaining 18% of taxa 
get_taxa_unique(ps2, "Genus") # 100 genera

ps1<-filter_taxa(ps, function(x) mean(x) >1, TRUE)
ntaxa(ps1) # 312, retaining 24% of taxa
get_taxa_unique(ps1, "Genus") # 130 genera

# export tables with low abundance taxa removed
otu = as(otu_table(ps1), "matrix")
taxon = as(tax_table(ps1), "matrix")
metadata = as(sample_data(ps1), "matrix")
write.table(otu,"silva_nochloronomito_otu_table_ps1.txt",sep="\t",col.names=NA)
write.table(taxon,"silva_nochloronomito_taxa_table_ps1.txt",sep="\t",col.names=NA)

# export ASV table as relative abundance
ps_ra<-transform_sample_counts(ps1, function(OTU) OTU/sum(OTU))
otu_ra = as(otu_table(ps_ra), "matrix")
write.table(otu_ra,"silva_nochloronomito_otu_table_ps1_RA.txt",sep="\t",col.names=NA)

```
# ALPHA DIVERSITY

```{r}
# Clear data and read in tables with all ASVs
otu <- read.table("silva_nochloronomito_otu_table.txt",sep="\t",header=TRUE, row.names=1)
taxon <- read.table("silva_nochloronomito_taxa_table.txt",sep="\t",header=TRUE,row.names=1)
samples<-read.table("metadata.txt",sep="\t",header=T,row.names=1)
OTU = otu_table(otu, taxa_are_rows=FALSE)
taxon<-as.matrix(taxon)
TAX = tax_table(taxon)
sampledata = sample_data(samples)
ps <- phyloseq(otu_table(otu, taxa_are_rows=FALSE), 
               sample_data(samples), 
               tax_table(taxon))
ps # 1321 taxa and 34 samples

# calculate alpha diversity measures with phyloseq wrapper
alpha.diversity <- estimate_richness(ps, measures = c("Observed","Shannon","Simpson"))

# combine alpha diversity measures with sample metadata
alpha.diversity <- tibble::rownames_to_column(alpha.diversity, "sample")
# fix sample names
alpha.diversity$sample <- gsub("\\.","-",alpha.diversity$sample)
samples <- tibble::rownames_to_column(samples, "sample")
alpha_meta <- merge(alpha.diversity, samples, by="sample", all=TRUE)

#set aesthetics 
theme_set(theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank()))
cols<-c("DD"="#D55E00","HD"="#0072B2")
alpha_meta$Health<-factor(alpha_meta$Health,levels=c("HD","DD"))

# plot box and whiskers with points on top
obs <- ggplot(alpha_meta,aes(x=Health,y=Observed))+
  geom_boxplot(outlier.shape=NA)+
  geom_jitter(aes(color=Health))+
  theme(legend.position="none")+
  theme(axis.title.x=element_blank())+
  ylab("Observed Richness")+
  theme(text=element_text(size=14,face="bold"))+
  scale_color_manual(values=cols)
obs

shan <- ggplot(alpha_meta,aes(x=Health,y=Shannon))+
  geom_boxplot(outlier.shape=NA)+
  geom_jitter(aes(color=Health))+
  theme(legend.position="none")+
  theme(axis.title.x=element_blank())+
  theme(text=element_text(size=14,face="bold"))+
  scale_color_manual(values=cols)
shan

simp <- ggplot(alpha_meta,aes(x=Health,y=Simpson))+
  geom_boxplot(outlier.shape=NA)+
  geom_jitter(aes(color=Health))+
  theme(legend.position="none")+
  theme(axis.title.x=element_blank())+
  theme(text=element_text(size=14,face="bold"))+
  scale_color_manual(values=cols)
simp

# plot 3 panels
pdf("AlphaDiversity.pdf", bg ="white", width=8, height=4)
plot_grid(obs,shan,simp,ncol=3)
dev.off()


# Test statistical differences in alpha diversity
# First test for normality using Shapiro test. If the value of p is equal to or less than 0.05, then the hypothesis of normality will be rejected. 
shapiro.test(alpha.diversity$Observed) # p-value = 4.89e-09
shapiro.test(alpha.diversity$Shannon) # p-value = 0.0029
shapiro.test(alpha.diversity$Simpson) # p-value = 0.0004532
# normality is rejected for all

# ANOVA on ranked data for non-normally distributed diversity measures - Observed, Shannon, Simpson
anova1 <- aov(rank(Observed) ~ Health, alpha_meta) 
summary(anova1) 
TukeyHSD(anova1)
# not significant

anova2 <- aov(rank(Shannon) ~ Health, alpha_meta) 
summary(anova2) 
TukeyHSD(anova2)
# not significant

anova3 <- aov(rank(Simpson) ~ Health, alpha_meta) 
summary(anova3) 
TukeyHSD(anova3)
# not significant
```

# BETA DIVERSITY

```{r}
# Clear data and read in tables with filtered ASVs (low abundance taxa removed)
otu <- read.table("silva_nochloronomito_otu_table_ps1.txt",sep="\t",header=TRUE, row.names=1)
taxon <- read.table("silva_nochloronomito_taxa_table_ps1.txt",sep="\t",header=TRUE,row.names=1)
samples<-read.table("metadata.txt",sep="\t",header=T,row.names=1)
OTU = otu_table(otu, taxa_are_rows=FALSE)
taxon<-as.matrix(taxon)
TAX = tax_table(taxon)
sampledata = sample_data(samples)
ps1 <- phyloseq(otu_table(otu, taxa_are_rows=FALSE), 
               sample_data(samples), 
               tax_table(taxon))
ps1 # 312 taxa and 34 samples


# CALCULATE AITCHISON DISTANCE
# First, replace 0 values with an estimate (because normalization is taking log, can't have 0)
# Also transposing here, need samples as rows
d.czm <- cmultRepl(t(otu), method="CZM", label=0, z.warning=1)
# Perform the center-log-ratio (CLR) transformation 
d.clr <- codaSeq.clr(d.czm)
# transpose matrix of CLR transformed data for ordination and dendrogram
E.clr <- t(d.clr)
# plot compositional PCA biplot (perform a singular value decomposition)
d.pcx <- prcomp(E.clr)
# calculate percent variance explained for the axis labels
pc1 <- round(d.pcx$sdev[1]^2/sum(d.pcx$sdev^2),2)
pc2 <- round(d.pcx$sdev[2]^2/sum(d.pcx$sdev^2),2)
xlab <- paste("PC1: ", pc1, sep="")
ylab <- paste("PC2: ", pc2, sep="")
# informational plots, not for publication
#biplot(d.pcx, cex=c(0.6,0.4), var.axes=F,scale=1, xlab=xlab, ylab=ylab)
screeplot(d.pcx)

# replot PCA with ggplot2 (showing samples only)
# first, create a dataframe with the PCA values
df_out <- as.data.frame(d.pcx$x)
# create a single dataframe for plotting in case samples are ordered differently in the pcx dataframe than in the sample metadata - in this dataset, it looks like they are already in the same order but samples names that start with numbers instead of letters can be an issue
# step 1 make rownames a column for that I can merge by sample name
samples2<-rownames_to_column(samples, var = "sample")
df_out2<-rownames_to_column(df_out, var = "sample")
single_df<-merge(samples2,df_out2, by="sample")

#set aesthetics 
theme_set(theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank()))
cols<-c("DD"="#D55E00","HD"="#0072B2")
single_df$Health<-factor(single_df$Health,levels=c("HD","DD"))

pdf("PCA.pdf",bg ="white",width = 8)
p<-ggplot(single_df,aes(x=PC1,y=PC2,color=Health))
p<-p+geom_point(size=2)+
  theme(axis.title = element_text(size=14))+
  theme(axis.text=element_text(size=12))+
  theme(legend.title = element_text(size=14))+
  theme(legend.text = element_text(size=12))+
  theme(strip.text.x = element_text(size=12))+
  scale_color_manual(values=cols)+
  guides(fill = guide_legend(override.aes=list(shape=21)))
p + labs(x=xlab, y=ylab)
dev.off()

####### Use phyloseq/vegan to perform PERMANOVA
# permanova between groups using Aitchison distance
dist.clr <- dist(E.clr)
perm<-adonis2(dist.clr~Health*Treatment,as(sample_data(ps1),"data.frame"),by="terms") # here "Treatment" was applied after microbiome sample was collected
print(perm)

# Permutation test for adonis under reduced model
# Terms added sequentially (first to last)
# Permutation: free
# Number of permutations: 999

# adonis2(formula = dist.clr ~ Health * Treatment, data = as(sample_data(ps1), "data.frame"), by = "terms")
#                  Df SumOfSqs      R2      F Pr(>F)  
# Health            1     1952 0.04900 1.6690  0.022 *
# Treatment         1     1536 0.03857 1.3135  0.092 .
# Health:Treatment  1     1258 0.03157 1.0752  0.295  
# Residual         30    35088 0.88086                
# Total            33    39834 1.00000
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

perm2<-adonis2(dist.clr~Health,as(sample_data(ps1),"data.frame"),by="terms")
print(perm2)

# Permutation test for adonis under reduced model
# Terms added sequentially (first to last)
# Permutation: free
# Number of permutations: 999

# adonis2(formula = dist.clr ~ Health, data = as(sample_data(ps1), "data.frame"), by = "terms")
#          Df SumOfSqs      R2     F Pr(>F)  
# Health    1     1952 0.049 1.6489  0.024 *
# Residual 32    37882 0.951                
# Total    33    39834 1.000                
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

# Health condition of sample (DD or HD) explains 5% of variation in the microbial community

```

# Examine microbiome composition

```{r}
# Create phyloseq object with relative abundance of filtered ASVs (low abundance taxa removed)
otu <- read.table("silva_nochloronomito_otu_table_ps1_RA.txt",sep="\t",header=TRUE, row.names=1)
taxon <- read.table("silva_nochloronomito_taxa_table_ps1.txt",sep="\t",header=TRUE,row.names=1)
samples<-read.table("metadata.txt",sep="\t",header=T,row.names=1)
OTU = otu_table(otu, taxa_are_rows=FALSE)
taxon<-as.matrix(taxon)
TAX = tax_table(taxon)
sampledata = sample_data(samples)
ps_ra <- phyloseq(otu_table(otu, taxa_are_rows=FALSE), 
               sample_data(samples), 
               tax_table(taxon))
ps_ra # 312 taxa and 34 samples

get_taxa_unique(ps_ra, "Genus") # 130 genera
get_taxa_unique(ps_ra, "Family") # 73 families
get_taxa_unique(ps_ra, "Order") # 43 orders

# Plotting most abundant families - first collapse ASVs into families then reshape data for plotting
ps_ra_f = tax_glom(ps_ra, "Family")
ps_ra_f #   families
otu_f = as(otu_table(ps_ra_f), "matrix")
taxon_f = as(tax_table(ps_ra_f), "matrix")
fam<-t(otu_f)
fam <- as.data.frame(fam)
fam <- tibble::rownames_to_column(fam, "ASV")
taxon_f <- as.data.frame(taxon_f)
taxon_f <- tibble::rownames_to_column(taxon_f, "ASV")
fam_tax <- merge(taxon_f, fam, by="ASV", all=FALSE)
fam_tax$Genus<-NULL
fam_tax$ASV<-NULL
# Order by total relative abundance and choose top 25 families
fam_tax <- fam_tax %>% mutate("total" = rowSums((fam_tax[,6:39]), na.rm = TRUE))  
fam_tax <- fam_tax[order(fam_tax$total, decreasing = TRUE),]
fam25 <- fam_tax[1:25,]
fam25$total<-NULL
# write out this table to sort the families by taxonomy
fam_long<-melt(fam25,value.name="proportion",variable.name="sample",id.vars=c("Kingdom","Phylum","Class","Order","Family"))
# Merge with metadata for plotting aesthetics
samples <- as.data.frame(samples)
samples <- tibble::rownames_to_column(samples, "sample")
famplot <- merge(fam_long,samples,by="sample", all=TRUE)
# how many classes or orders in the top 25 families?
unique(famplot$Class) # 10 classes:"Gammaproteobacteria","Alphaproteobacteria","Campylobacteria","Bacteroidia","Cyanobacteriia","Clostridia","Blastocatellia","Verrucomicrobiia","Polyangiia","Desulfovibrionia" 
unique(famplot$Family)
#  [1] "Vibrionaceae"            "Pseudoalteromonadaceae"  "Paracoccaceae"           "Rhodospirillales"        "Arcobacteraceae"        
# [6] "Flavobacteriaceae"       "Cyanobiaceae"            "Alteromonadaceae"        "Nitrincolaceae"          "Colwelliaceae"          
# [11] "Clostridiaceae"          "Blastocatellaceae"       "Ga0077536"               "Saccharospirillaceae"    "Rhizobiaceae"           
# [16] "Shewanellaceae"          "Rubritaleaceae"          "Sandaracinaceae"         "Clade I"                 "SAR116 clade"           
# [21] "Moritellaceae"           "AEGEAN-169 marine group" "Cyclobacteriaceae"       "Alcaligenaceae"          "Desulfovibrionaceae" 

# set aesthetics
theme_set(theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank()))
famplot$Health<-factor(famplot$Health,levels=c("HD","DD"))
labels <-c(HD = "Apparently Healthy",DD = "Disease Lesion")
mycols<-c("Alphaproteobacteria"="#CC6677","Bacteroidia"="#88CCEE","Blastocatellia"="#999933","Campylobacteria"="#882255","Clostridia"="#000000","Cyanobacteriia"="#44AA99","Desulfovibrionia"="#DDCC77","Gammaproteobacteria"="#AA4499","Polyangiia"="#332288","Verrucomicrobiia"="#117733")

# order families alphabetically under class, first = bottom of plot
famplot$Family<-factor(famplot$Family, levels=c("Rubritaleaceae","Sandaracinaceae","Saccharospirillaceae","Nitrincolaceae","Ga0077536","Vibrionaceae","Shewanellaceae","Pseudoalteromonadaceae","Moritellaceae","Colwelliaceae","Alteromonadaceae","Alcaligenaceae","Desulfovibrionaceae","Cyanobiaceae","Clostridiaceae","Arcobacteraceae","Blastocatellaceae","Flavobacteriaceae","Cyclobacteriaceae","Rhodospirillales","AEGEAN-169 marine group","Paracoccaceae","SAR116 clade","Clade I","Rhizobiaceae"))

pdf("bubble_25families.pdf",bg ="white",width=8)
bubble <- ggplot(famplot, aes(x=sample,y=Family,size=proportion,color=Class))+
  geom_point(alpha=0.8)+ 
  scale_size(range = c(0,10), name="Relative abundance")+
  scale_color_manual(values=mycols)+
  facet_grid(Health ~ ., scales="free", space="free", labeller=labeller(Health=labels))+
  theme(axis.text.x=element_blank())+  
  theme(axis.text.y=element_text(size=10))+
  theme(axis.title.y=element_blank())+
  xlab("Colony")+
  theme(legend.text = element_text(size=10))
bubble  
dev.off()


```

# Differentially abundant taxa


```{r}

# Clear data and read in tables with all ASVs and the ASV sequence has been replaced with a number
otu <- read.table("silva_nochloronomito_otu_table_corncob.txt",sep="\t",header=TRUE, row.names=1)
taxon <- read.table("silva_nochloronomito_taxa_table_corncob.txt",sep="\t",header=TRUE,row.names=1)
samples<-read.table("metadata.txt",sep="\t",header=T,row.names=1)
OTU = otu_table(otu, taxa_are_rows=FALSE)
taxon<-as.matrix(taxon)
TAX = tax_table(taxon)
sampledata = sample_data(samples)
ps <- phyloseq(otu_table(otu, taxa_are_rows=FALSE), 
               sample_data(samples), 
               tax_table(taxon))
ps # 1321 taxa and 34 samples

set.seed(1)
tissue.da <- differentialTest(formula = ~ Percent_healthy_remaining_1month, 
                                 phi.formula = ~ Percent_healthy_remaining_1month,
                                 formula_null = ~ 1,
                                 phi.formula_null = ~ Percent_healthy_remaining_1month,
                                 test = "LRT", boot = FALSE,
                                 data = ps,
                                 fdr_cutoff = 0.05)

summary(tissue.da)
tissue.da
tissue.da$significant_taxa # n = 28
# [1] "ASV329"  "ASV374"  "ASV395"  "ASV438"  "ASV498"  "ASV821"  "ASV992"  "ASV993"  "ASV994"  "ASV995"  "ASV996"  "ASV997"  "ASV998" 
# [14] "ASV1151" "ASV1152" "ASV1167" "ASV1259" "ASV1260" "ASV1261" "ASV1262" "ASV1263" "ASV1264" "ASV1265" "ASV1266" "ASV1267" "ASV1268"
# [27] "ASV1294" "ASV1295"
plot(tissue.da, c("Genus"))


sigtaxto_df <- function(tissue.da, ps) {
  df.out <- c()
  for (i in 1:length(tissue.da$significant_models)) { #from 1 to the number of significant taxa
    df.out = rbind(df.out, tissue.da$significant_models[[i]]$coefficients[2, 1:4]) #bind the significant taxa with coefficient, std. error, t value, Pr. coefficients[2,] is the DA model output.
  }
  asv = as.vector(tissue.da$significant_taxa) #get significant ASVs
  Class = as.vector(otu_to_taxonomy(data = ps, tissue.da$significant_taxa, level = c("Class")))
  Order = as.vector(otu_to_taxonomy(data = ps, tissue.da$significant_taxa, level = c("Order")))
  Family = as.vector(otu_to_taxonomy(data = ps, tissue.da$significant_taxa, level = c("Family")))
  Genus = as.vector(otu_to_taxonomy(data = ps, tissue.da$significant_taxa, level = c("Genus")))
  
  df.bind <- as.data.frame(cbind(df.out, asv, Class, Order, Family, Genus))
  df.bind$asvgenus <- paste0(df.bind$Genus, " (", df.bind$asv, ")")
  colnames(df.bind) <- c("Estimate", "StdError", "tvalue", "Pr", "asv", "Class", "Order", "Family", "Genus", "asvtaxa")
  df.bind$Estimate = as.numeric(as.character(df.bind$Estimate)) #estimate column should be numeric 
  df.bind$StdError = as.numeric(as.character(df.bind$StdError)) #stderror column should be numeric 
  
  return(df.bind)
}

df <- sigtaxto_df(tissue.da, ps) #make a df of the corncob output and the taxonomy

### Save these data so you don't have to re-run the model ###
write.table(df, "DA-taxa-tissue.txt", sep="\t",row.names = FALSE)





```


