---
title: "Cnat2-18_field_trials"
author: "J. Meyer"
date: "2026-07-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries

```{r}
library(dada2)
library(CoDaSeq)
library(ggplot2)
library(phyloseq)
library(vegan)
library(pairwiseAdonis)
library(corncob)
library(knitr)
library(dplyr)
library(tibble)
library(reshape2)
library(cowplot)
writeLines(capture.output(sessionInfo()), "sessionInfo.txt")
```


## Quality-filter the sequencing reads and create Amplicon Sequence Variant (ASV) tables with DADA2
Put unjoined R1 and R2 fastq files, with adaptors and primers previously removed with cutadapt into a directory for DADA2. Here, our forward and reverse fastq filenames have format: SAMPLENAME_R1_cut.fastq.gz and SAMPLENAME_R2_cut.fastq.gz


```{r}
path <- "~/Documents/CNAT2-18_fieldtrials/cutadapt"
list.files(path)
fnFs <- sort(list.files(path, pattern="_R1_cut.fastq.gz", full.names = TRUE))
fnRs <- sort(list.files(path, pattern="_R2_cut.fastq.gz", full.names = TRUE))
sample.names <- sapply(strsplit(basename(fnFs), "_"), `[`, 1)
# Perform filtering and trimming
filt_path <- file.path(path, "filtered") 
filtFs <- file.path(filt_path, paste0(sample.names, "_F_filt.fastq.gz"))
filtRs <- file.path(filt_path, paste0(sample.names, "_R_filt.fastq.gz"))
out <- filterAndTrim(fnFs, filtFs, fnRs, filtRs, truncLen=c(150,150),
                     maxN=0, maxEE=c(2,2), truncQ=2, rm.phix=TRUE,
                     compress=TRUE, multithread=TRUE)
head(out)
# Learn the Error Rates, it TAKES TIME!
errF <- learnErrors(filtFs, multithread=TRUE)
errR <- learnErrors(filtRs, multithread=TRUE)
plotErrors(errF, nominalQ=TRUE)
# Dereplicate the filtered fastq files
derepFs <- derepFastq(filtFs, verbose=TRUE)
derepRs <- derepFastq(filtRs, verbose=TRUE)
names(derepFs) <- sample.names
names(derepRs) <- sample.names
# Infer the sequence variants in each sample
dadaFs <- dada(derepFs, err=errF, multithread=TRUE)
dadaRs <- dada(derepRs, err=errR, multithread=TRUE)
# Inspecting the dada-class object returned by dada:
dadaFs[[1]]
# Merge the denoised forward and reverse reads:
mergers <- mergePairs(dadaFs, derepFs, dadaRs, derepRs, verbose=TRUE)
# Inspect the merger data.frame from the first sample
head(mergers[[1]])
# Construct sequence table
seqtab <- makeSequenceTable(mergers)
dim(seqtab)
#Remove chimeric sequences:
seqtab.nochim <- removeBimeraDenovo(seqtab, method="consensus", multithread=TRUE, verbose=TRUE)
dim(seqtab.nochim)
sum(seqtab.nochim)/sum(seqtab)
# Inspect distribution of sequence lengths
table(nchar(getSequences(seqtab)))
# Track reads through the pipeline
getN <- function(x) sum(getUniques(x))
track <- cbind(out, sapply(dadaFs, getN), sapply(mergers, getN), rowSums(seqtab))
colnames(track) <- c("input", "filtered", "denoised", "merged", "tabled")
rownames(track) <- sample.names
head(track)
write.table(track, "dada_read_stats.txt",sep="\t",col.names=NA)
saveRDS(seqtab.nochim, "~/Documents/CNAT2-18_fieldtrials/seqtab.rds") 
```


## Assign taxonomy in DADA2

Make sure the taxonomy reference database is in your working directory. Keep the database file gzipped. Adjust path name below. This step is very time consuming.

When taxonomy assignment is complete, we will use base R and phyloseq to clean up the taxonomy table. First, we will replace NAs and empty cells with the lowest taxonomy classification available. Second, we will use phyloseq to remove reads that are classified as Eukaryotes or unclassified at the domain level (ie, we are keeping only Bacteria and Archaea because that is what our primers target).

```{r}
taxa <- assignTaxonomy(seqtab.nochim, "~/Documents/silva_nr99_v138.2_toGenus_trainset.fa.gz", multithread=TRUE)
# FIX the NAs in the taxa table
taxon <- as.data.frame(taxa,stringsAsFactors=FALSE)
taxon$Phylum[is.na(taxon$Phylum)] <- taxon$Kingdom[is.na(taxon$Phylum)]
taxon$Class[is.na(taxon$Class)] <- taxon$Phylum[is.na(taxon$Class)]
taxon$Order[is.na(taxon$Order)] <- taxon$Class[is.na(taxon$Order)]
taxon$Family[is.na(taxon$Family)] <- taxon$Order[is.na(taxon$Family)]
taxon$Genus[is.na(taxon$Genus)] <- taxon$Family[is.na(taxon$Genus)]
write.table(taxon,"silva_taxa_table.txt",sep="\t",col.names=NA)
write.table(seqtab.nochim, "silva_otu_table.txt",sep="\t",col.names=NA)

# Create phyloseq object from otu and taxonomy tables from dada2, along with the sample metadata.
otu <- read.table("silva_otu_table.txt",sep="\t",header=TRUE, row.names=1)
taxon <- read.table("silva_taxa_table.txt",sep="\t",header=TRUE,row.names=1)
samples<-read.table("metadata.txt",sep="\t",header=T,row.names=1)
OTU = otu_table(otu, taxa_are_rows=FALSE)
taxon<-as.matrix(taxon)
TAX = tax_table(taxon)
sampledata = sample_data(samples)
ps <- phyloseq(otu_table(otu, taxa_are_rows=FALSE), 
               sample_data(samples), 
               tax_table(taxon))
ps #2048 taxa and 58 samples

# remove chloroplasts and mitochondria and Eukaryota
get_taxa_unique(ps, "Family")
get_taxa_unique(ps, "Order")
get_taxa_unique(ps, "Kingdom")
ps <- subset_taxa(ps, Family !="Mitochondria")
ps <- subset_taxa(ps, Order !="Chloroplast")
ps <- subset_taxa(ps, Kingdom !="Eukaryota")
ps <- subset_taxa(ps, Kingdom !="NA")
get_taxa_unique(ps, "Family")
get_taxa_unique(ps, "Order") 
get_taxa_unique(ps, "Kingdom")
ps # 1984 taxa and 58 samples

# Now export cleaned otu and taxa tables from phyloseq for future reference
otu = as(otu_table(ps), "matrix")
taxon = as(tax_table(ps), "matrix")
metadata = as(sample_data(ps), "matrix")
write.table(otu,"silva_nochloronomito_otu_table.txt",sep="\t",col.names=NA)
write.table(taxon,"silva_nochloronomito_taxa_table.txt",sep="\t",col.names=NA)

# explore data
length(get_taxa_unique(ps, "Genus")) # 529 genera
length(get_taxa_unique(ps, "Family")) # 284 families
length(get_taxa_unique(ps, "Order")) # 175 orders
length(get_taxa_unique(ps, "Class")) # 78 classes
length(get_taxa_unique(ps, "Phylum")) # 36 phyla

ps5<-filter_taxa(ps, function(x) mean(x) >5, TRUE)
ntaxa(ps5) #199 - this retains 10% of original ASVs
get_taxa_unique(ps5, "Genus") # 84 genera

ps2<-filter_taxa(ps, function(x) mean(x) >2, TRUE)
ntaxa(ps2) #325 
get_taxa_unique(ps2, "Genus") #129 genera

ps1<-filter_taxa(ps, function(x) mean(x) >1, TRUE)
ntaxa(ps1) #441, retaining 22% of taxa
get_taxa_unique(ps1, "Genus") #160 genera


# export tables with low abundance taxa removed
otu = as(otu_table(ps1), "matrix")
taxon = as(tax_table(ps1), "matrix")
metadata = as(sample_data(ps1), "matrix")
write.table(otu,"silva_nochloronomito_otu_table_ps1.txt",sep="\t",col.names=NA)
write.table(taxon,"silva_nochloronomito_taxa_table_ps1.txt",sep="\t",col.names=NA)

# export ASV table as relative abundance
ps_ra<-transform_sample_counts(ps1, function(OTU) OTU/sum(OTU))
otu_ra = as(otu_table(ps_ra), "matrix")
write.table(otu_ra,"silva_nochloronomito_otu_table_ps1_RA.txt",sep="\t",col.names=NA)

```
# BETA DIVERSITY

```{r}
# Create phyloseq object with filtered ASVs (low abundance taxa removed)
otu <- read.table("silva_nochloronomito_otu_table_ps1.txt",sep="\t",header=TRUE, row.names=1)
taxon <- read.table("silva_nochloronomito_taxa_table_ps1.txt",sep="\t",header=TRUE,row.names=1)
samples<-read.table("metadata.txt",sep="\t",header=T,row.names=1)
OTU = otu_table(otu, taxa_are_rows=FALSE)
taxon<-as.matrix(taxon)
TAX = tax_table(taxon)
sampledata = sample_data(samples)
ps1 <- phyloseq(otu_table(otu, taxa_are_rows=FALSE), 
               sample_data(samples), 
               tax_table(taxon))
ps1 # 441 taxa and 58 samples

# CALCULATE AITCHISON DISTANCE
# First, replace 0 values with an estimate (because normalization is taking log, can't have 0)
# Also transposing here, need samples as rows
d.czm <- cmultRepl(t(otu), method="CZM", label=0, z.warning=1)
# Perform the center-log-ratio (CLR) transformation 
d.clr <- codaSeq.clr(d.czm)
# transpose matrix of CLR transformed data for ordination and dendrogram
E.clr <- t(d.clr)
# plot compositional PCA biplot (perform a singular value decomposition)
d.pcx <- prcomp(E.clr)
# calculate percent variance explained for the axis labels
pc1 <- round(d.pcx$sdev[1]^2/sum(d.pcx$sdev^2),2)
pc2 <- round(d.pcx$sdev[2]^2/sum(d.pcx$sdev^2),2)
xlab <- paste("PC1: ", pc1, sep="")
ylab <- paste("PC2: ", pc2, sep="")
# informational plots, not for publication
#biplot(d.pcx, cex=c(0.6,0.4), var.axes=F,scale=1, xlab=xlab, ylab=ylab)
screeplot(d.pcx)

# replot PCA with ggplot2 (showing samples only)
# first, create a dataframe with the PCA values
df_out <- as.data.frame(d.pcx$x)
# create a single dataframe for plotting in case samples are ordered differently in the pcx dataframe than in the sample metadata - in this dataset, it looks like they are already in the same order but samples names that start with numbers instead of letters can be an issue
# step 1 make rownames a column for that I can merge by sample name
samples2<-rownames_to_column(samples, var = "sample")
df_out2<-rownames_to_column(df_out, var = "sample")
single_df<-merge(samples2,df_out2, by="sample")

#set aesthetics 
theme_set(theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank()))
cols<-c("DD"="#D55E00","HD"="#56B4E9","HH"="#0072B2")

p<-ggplot(single_df,aes(x=PC1,y=PC2,color=Health,shape=Month))
p<-p+geom_point(size=2)+
  theme(axis.title = element_text(size=14))+
  theme(axis.text=element_text(size=12))+
  theme(legend.title = element_text(size=14))+
  theme(legend.text = element_text(size=12))+
  theme(strip.text.x = element_text(size=12))+
  scale_color_manual(values=cols)+
  guides(fill = guide_legend(override.aes=list(shape=21)))
p + labs(x=xlab, y=ylab)

####### Use phyloseq/vegan to perform PERMANOVA
# permanova between groups using Aitchison distance
dist.clr <- dist(E.clr)
perm<-adonis2(dist.clr~Health*Treatment,as(sample_data(ps1),"data.frame"),by="terms")
print(perm)
# Microbial community variation is explained by Health (R2 = 0.07083, p < 0.001), but not by treatment

perm<-adonis2(dist.clr~Health*Month,as(sample_data(ps1),"data.frame"),by="terms")
print(perm)
# Microbial community variation is explained by Health (R2 = 0.07083, p < 0.001), by Month (R2 = 0.08554, p < 0.001), and by the interaction of Health and Month (R2 = 0.04778, p = 0.001)

perm2<-pairwise.adonis2(dist.clr~Health,as(sample_data(ps1),"data.frame"),by = "terms")
print(perm2)
# DD sign diff from HD, no other sign diff

perm3<-pairwise.adonis2(dist.clr~Health/Month,as(sample_data(ps1),"data.frame"), strata='Month')
print(perm3)
# DD sign diff from HD, no other sign diff
```

# Examine microbiome composition

```{r}
# Create phyloseq object with relative abundance of filtered ASVs (low abundance taxa removed)
otu <- read.table("silva_nochloronomito_otu_table_ps1_RA.txt",sep="\t",header=TRUE, row.names=1)
taxon <- read.table("silva_nochloronomito_taxa_table_ps1.txt",sep="\t",header=TRUE,row.names=1)
samples<-read.table("metadata.txt",sep="\t",header=T,row.names=1)
OTU = otu_table(otu, taxa_are_rows=FALSE)
taxon<-as.matrix(taxon)
TAX = tax_table(taxon)
sampledata = sample_data(samples)
ps_ra <- phyloseq(otu_table(otu, taxa_are_rows=FALSE), 
               sample_data(samples), 
               tax_table(taxon))
ps_ra # 441 taxa and 58 samples

get_taxa_unique(ps_ra, "Genus") #160 genera
get_taxa_unique(ps_ra, "Family") #92 families
get_taxa_unique(ps_ra, "Order") #51 orders

# Plotting most abundant families - first collapse ASVs into families then reshape data for plotting
ps_ra_f = tax_glom(ps_ra, "Family")
ps_ra_f # 92 families
otu_f = as(otu_table(ps_ra_f), "matrix")
taxon_f = as(tax_table(ps_ra_f), "matrix")
fam<-t(otu_f)
fam <- as.data.frame(fam)
fam <- tibble::rownames_to_column(fam, "ASV")
taxon_f <- as.data.frame(taxon_f)
taxon_f <- tibble::rownames_to_column(taxon_f, "ASV")
fam_tax <- merge(taxon_f, fam, by="ASV", all=FALSE)
fam_tax$Genus<-NULL
fam_tax$ASV<-NULL
# Order by total relative abundance and choose top 25 families
fam_tax <- fam_tax %>% mutate("total" = rowSums((fam_tax[,6:63]), na.rm = TRUE))  
fam_tax <- fam_tax[order(fam_tax$total, decreasing = TRUE),]
fam25 <- fam_tax[1:25,]
fam25$total<-NULL
fam_long<-melt(fam25,value.name="proportion",variable.name="sample",id.vars=c("Kingdom","Phylum","Class","Order","Family"))
# Merge with metadata for plotting aesthetics
samples <- as.data.frame(samples)
samples <- tibble::rownames_to_column(samples, "sample")
famplot <- merge(fam_long,samples,by="sample", all=TRUE)
# how many classes or orders in the top 25 families?
unique(famplot$Class) # 9 classes:"Gammaproteobacteria","Alphaproteobacteria","Bacteroidia","Campylobacteria","Cyanobacteriia","Verrucomicrobiia","Clostridia","Blastocatellia","Acidimicrobiia"
unique(famplot$Family)
# [1] "Vibrionaceae"            "Pseudoalteromonadaceae"  "Paracoccaceae"           "Flavobacteriaceae"       "Alteromonadaceae"        "Arcobacteraceae"        
# [7] "Rhodospirillales"        "Cyanobiaceae"            "Nitrincolaceae"          "Colwelliaceae"           "Saccharospirillaceae"    "Rubritaleaceae"         
# [13] "Rhizobiaceae"            "Ga0077536"               "Clostridiaceae"          "Blastocatellaceae"       "Shewanellaceae"          "Actinomarinaceae"       
# [19] "Alcaligenaceae"          "Cyclobacteriaceae"       "AEGEAN-169 marine group" "Clade I"                 "Sphingomonadaceae"       "Clade III"              
# [25] "SAR116 clade"

# set aesthetics
theme_set(theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank()))
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7","#000000")
# order families alphabetically under class, first = bottom of plot
famplot$Family<-factor(famplot$Family, levels=c("Rubritaleaceae","Vibrionaceae","Shewanellaceae","Saccharospirillaceae","Pseudoalteromonadaceae","Nitrincolaceae","Ga0077536","Colwelliaceae","Alteromonadaceae","Alcaligenaceae","Cyanobiaceae","Clostridiaceae","Arcobacteraceae","Blastocatellaceae","Flavobacteriaceae","Cyclobacteriaceae","Sphingomonadaceae","SAR116 clade","Rhodospirillales","Rhizobiaceae","Paracoccaceae","Clade III","Clade I","AEGEAN-169 marine group","Actinomarinaceae"))

#pdf("bubble_25families_withHH.pdf",bg ="white",width=11)
bubble <- ggplot(famplot, aes(x=sample,y=Family,size=proportion,color=Class))+
  geom_point(alpha=0.8)+ 
  scale_size(range = c(0,10), name="Relative abundance")+
  scale_color_manual(values=cbPalette)+
  facet_grid(Health ~ Month, scales="free", space="free")+
  theme(axis.text.x=element_blank())+  
  theme(axis.text.y=element_text(size=10))+
  theme(axis.title.x=element_blank())+
  theme(axis.title.y=element_blank())+
  theme(legend.text = element_text(size=10))
bubble  
#dev.off()


```

### Since we only have 4 HH samples (Sept only), let's re-analyze with those samples removed.


```{r}
# Create phyloseq object with filtered ASVs (low abundance taxa removed)
otu <- read.table("silva_nochloronomito_otu_table_ps1.txt",sep="\t",header=TRUE, row.names=1)
taxon <- read.table("silva_nochloronomito_taxa_table_ps1.txt",sep="\t",header=TRUE,row.names=1)
samples<-read.table("metadata.txt",sep="\t",header=T,row.names=1)
OTU = otu_table(otu, taxa_are_rows=FALSE)
taxon<-as.matrix(taxon)
TAX = tax_table(taxon)
sampledata = sample_data(samples)
ps1 <- phyloseq(otu_table(otu, taxa_are_rows=FALSE), 
               sample_data(samples), 
               tax_table(taxon))
ps1 # 441 taxa and 58 samples

#remove 4 HH samples
ps1_DD_HD = subset_samples(ps1, Health != "HH")
ps1_DD_HD # 441 taxa and 54 samples
#remove taxa only found in HH samples
ps1_DD_HD <- prune_taxa(taxa_sums(ps1_DD_HD) > 1, ps1_DD_HD)
ps1_DD_HD # 438 taxa and 54 samples

# export tables with HH samples removed
otu = as(otu_table(ps1_DD_HD), "matrix")
taxon = as(tax_table(ps1_DD_HD), "matrix")
metadata = as(sample_data(ps1_DD_HD), "matrix")
write.table(otu,"silva_nochloronomito_otu_table_ps1_DD_HD.txt",sep="\t",col.names=NA)
write.table(taxon,"silva_nochloronomito_taxa_table_ps1_DD_HD.txt",sep="\t",col.names=NA)
write.table(metadata,"metadata_ps1_DD_HD.txt",sep="\t",col.names=NA)

# export ASV table as relative abundance
ps_ra<-transform_sample_counts(ps1_DD_HD, function(OTU) OTU/sum(OTU))
otu_ra = as(otu_table(ps_ra), "matrix")
write.table(otu_ra,"silva_nochloronomito_otu_table_ps1_DD_HD_RA.txt",sep="\t",col.names=NA)

# clear data and read back in the dataset with 54 samples
otu <- read.table("silva_nochloronomito_otu_table_ps1_DD_HD.txt",sep="\t",header=TRUE, row.names=1)
taxon <- read.table("silva_nochloronomito_taxa_table_ps1_DD_HD.txt",sep="\t",header=TRUE,row.names=1)
samples<-read.table("metadata_ps1_DD_HD.txt",sep="\t",header=T,row.names=1)
OTU = otu_table(otu, taxa_are_rows=FALSE)
taxon<-as.matrix(taxon)
TAX = tax_table(taxon)
sampledata = sample_data(samples)
ps1_DD_HD <- phyloseq(otu_table(otu, taxa_are_rows=FALSE), 
               sample_data(samples), 
               tax_table(taxon))
ps1_DD_HD # 438 taxa and 54 samples


# CALCULATE AITCHISON DISTANCE
# First, replace 0 values with an estimate (because normalization is taking log, can't have 0)
# Also transposing here, need samples as rows
d.czm <- cmultRepl(t(otu), method="CZM", label=0, z.warning=1)
# Perform the center-log-ratio (CLR) transformation 
d.clr <- codaSeq.clr(d.czm)
# transpose matrix of CLR transformed data for ordination and dendrogram
E.clr <- t(d.clr)
# plot compositional PCA biplot (perform a singular value decomposition)
d.pcx <- prcomp(E.clr)
# calculate percent variance explained for the axis labels
pc1 <- round(d.pcx$sdev[1]^2/sum(d.pcx$sdev^2),2)
pc2 <- round(d.pcx$sdev[2]^2/sum(d.pcx$sdev^2),2)
xlab <- paste("PC1: ", pc1, sep="")
ylab <- paste("PC2: ", pc2, sep="")
# informational plots, not for publication
#biplot(d.pcx, cex=c(0.6,0.4), var.axes=F,scale=1, xlab=xlab, ylab=ylab)
screeplot(d.pcx)

# replot PCA with ggplot2 (showing samples only)
# first, create a dataframe with the PCA values
df_out <- as.data.frame(d.pcx$x)
# create a single dataframe for plotting in case samples are ordered differently in the pcx dataframe than in the sample metadata - in this dataset, it looks like they are already in the same order but samples names that start with numbers instead of letters can be an issue
# step 1 make rownames a column for that I can merge by sample name
samples2<-rownames_to_column(samples, var = "sample")
df_out2<-rownames_to_column(df_out, var = "sample")
single_df<-merge(samples2,df_out2, by="sample")

#set aesthetics 
theme_set(theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank()))
cols<-c("DD"="#D55E00","HD"="#0072B2")

pdf("PCA.pdf",bg ="white",width = 8.5)
p<-ggplot(single_df,aes(x=PC1,y=PC2,color=Health,shape=Month))
p<-p+geom_point(size=2)+
  theme(axis.title = element_text(size=14))+
  theme(axis.text=element_text(size=12))+
  theme(legend.title = element_text(size=14))+
  theme(legend.text = element_text(size=12))+
  theme(strip.text.x = element_text(size=12))+
  scale_color_manual(values=cols)+
  guides(fill = guide_legend(override.aes=list(shape=21)))
p + labs(x=xlab, y=ylab)
dev.off()

####### Use phyloseq/vegan to perform PERMANOVA
# permanova between groups using Aitchison distance
dist.clr <- dist(E.clr)
perm<-adonis2(dist.clr~Health*Treatment,as(sample_data(ps1_DD_HD),"data.frame"),by="terms")
print(perm)
# Microbial community variation is explained by Health (R2 = 0.05899, p < 0.001), and by treatment (R2 = 0.04695, p < 0.05)

perm2<-adonis2(dist.clr~Health*Month,as(sample_data(ps1_DD_HD),"data.frame"),by="terms")
print(perm2)
# Microbial community variation is explained by Health (R2 = 0.05899, p < 0.001), by Month (R2 = 0.08858, p < 0.001), and by the interaction of Health and Month (R2 = 0.05029, p < 0.001)

perm3<-pairwise.adonis2(dist.clr~Health/Month,as(sample_data(ps1_DD_HD),"data.frame"), strata='Month')
print(perm3)
# DD sign diff from HD when controlling for month (R2 = 0.19786, p < 0.001)
```
# Examine microbiome composition

```{r}
# Create phyloseq object with relative abundance of filtered ASVs (low abundance taxa removed)
otu <- read.table("silva_nochloronomito_otu_table_ps1_DD_HD_RA.txt",sep="\t",header=TRUE, row.names=1)
taxon <- read.table("silva_nochloronomito_taxa_table_ps1_DD_HD.txt",sep="\t",header=TRUE,row.names=1)
samples<-read.table("metadata_ps1_DD_HD.txt",sep="\t",header=T,row.names=1)
OTU = otu_table(otu, taxa_are_rows=FALSE)
taxon<-as.matrix(taxon)
TAX = tax_table(taxon)
sampledata = sample_data(samples)
ps_ra <- phyloseq(otu_table(otu, taxa_are_rows=FALSE), 
               sample_data(samples), 
               tax_table(taxon))
ps_ra # 438 taxa and 54 samples

get_taxa_unique(ps_ra, "Genus") #159 genera
get_taxa_unique(ps_ra, "Family") #92 families
get_taxa_unique(ps_ra, "Order") #51 orders

# Plotting most abundant families - first collapse ASVs into families then reshape data for plotting
ps_ra_f = tax_glom(ps_ra, "Family")
ps_ra_f # 92 families
otu_f = as(otu_table(ps_ra_f), "matrix")
taxon_f = as(tax_table(ps_ra_f), "matrix")
fam<-t(otu_f)
fam <- as.data.frame(fam)
fam <- tibble::rownames_to_column(fam, "ASV")
taxon_f <- as.data.frame(taxon_f)
taxon_f <- tibble::rownames_to_column(taxon_f, "ASV")
fam_tax <- merge(taxon_f, fam, by="ASV", all=FALSE)
fam_tax$Genus<-NULL
fam_tax$ASV<-NULL
# Order by total relative abundance and choose top 25 families
fam_tax <- fam_tax %>% mutate("total" = rowSums((fam_tax[,6:59]), na.rm = TRUE))  
fam_tax <- fam_tax[order(fam_tax$total, decreasing = TRUE),]
fam25 <- fam_tax[1:25,]
fam25$total<-NULL
# how much of the total relative abundance is represented in these 25 families?
fam26<- fam25 %>% bind_rows(summarise(.,across(where(is.numeric), sum),across(where(is.character), ~"Total")))
min(fam26[26,6:59]) # 0.87417
max(fam26[26,6:59]) # 0.9994597

fam_long<-melt(fam25,value.name="proportion",variable.name="sample",id.vars=c("Kingdom","Phylum","Class","Order","Family"))
# Merge with metadata for plotting aesthetics
samples <- as.data.frame(samples)
samples <- tibble::rownames_to_column(samples, "sample")
famplot <- merge(fam_long,samples,by="sample", all=TRUE)
# how many classes or orders in the top 25 families?
unique(famplot$Class) 
# [1] "Gammaproteobacteria" "Alphaproteobacteria" "Bacteroidia"         "Campylobacteria"     "Cyanobacteriia"      "Verrucomicrobiia"    "Clostridia"         
# [8] "Blastocatellia"      "Acidimicrobiia" 
unique(famplot$Family)
#  [1] "Vibrionaceae"            "Pseudoalteromonadaceae"  "Paracoccaceae"           "Flavobacteriaceae"       "Arcobacteraceae"         "Alteromonadaceae"       
# [7] "Cyanobiaceae"            "Nitrincolaceae"          "Rhodospirillales"        "Colwelliaceae"           "Saccharospirillaceae"    "Rubritaleaceae"         
# [13] "Rhizobiaceae"            "Ga0077536"               "Clostridiaceae"          "Blastocatellaceae"       "Actinomarinaceae"        "Alcaligenaceae"         
# [19] "Shewanellaceae"          "Cyclobacteriaceae"       "AEGEAN-169 marine group" "Clade I"                 "Sphingomonadaceae"       "Clade III"              
# [25] "SAR116 clade"   


# set aesthetics
theme_set(theme_bw()+theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank()))
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7","#000000")
# order families alphabetically under class, first = bottom of plot
famplot$Family<-factor(famplot$Family, levels=c("Rubritaleaceae","Vibrionaceae","Shewanellaceae","Saccharospirillaceae","Pseudoalteromonadaceae","Nitrincolaceae","Ga0077536","Colwelliaceae","Alteromonadaceae","Alcaligenaceae","Cyanobiaceae","Clostridiaceae","Arcobacteraceae","Blastocatellaceae","Flavobacteriaceae","Cyclobacteriaceae","Sphingomonadaceae","SAR116 clade","Rhodospirillales","Rhizobiaceae","Paracoccaceae","Clade III","Clade I","AEGEAN-169 marine group","Actinomarinaceae"))

pdf("bubble_25families.pdf",bg ="white",width=11)
bubble <- ggplot(famplot, aes(x=sample,y=Family,size=proportion,color=Class))+
  geom_point(alpha=0.8)+ 
  scale_size(range = c(0,10), name="Relative abundance")+
  scale_color_manual(values=cbPalette)+
  facet_grid(Health ~ Month, scales="free", space="free")+
  theme(axis.text.x=element_blank())+  
  theme(axis.text.y=element_text(size=10))+
  theme(axis.title.x=element_blank())+
  theme(axis.title.y=element_blank())+
  theme(legend.text = element_text(size=10))
bubble  
dev.off()


```



